---
title: "#3: Measuring Trees - R notebook"
# author: "Justinn J. Jones '24, GISP"
# date: "`r Sys.Date()`"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Howdy!

This R markdown notebook is the third step towards automating the process of measuring trees.

This notebook assumes the RMD resides directly in the "main directory" with a '\_input' folder housing the input files. We will create and use an 'output' folder for output files.

### 1.0 - Setup Notebook

#### Load packages

```{r load, echo = T, results = 'hide', warning=FALSE, message=FALSE}
library(TreeLS)
```

#### Setup the workspace

Verify the working directory.

```{r set wd}
wd <- getwd()
wd
```

Set input directory and list the files.

```{r input dir}
dir_input <- './_input'
list.files(dir_input)
```

Name the objects for later use.

```{r input vars}
las_in <- "subset_10_trees.las"
```

Create the 'output' directory. 

```{r output, results = 'hide'}
dir_out <- paste(wd, '/output/', sep = '')
ifelse(!dir.exists(file.path(dir_out)), dir.create(file.path(dir_out)), FALSE)
```

### 2.0 - Load Data

Read in the lidar point cloud and generate a summary of the dataset.

```{r load las, warning=FALSE}
tls <- readLAS(paste(dir_input, '/', las_in, sep = ''))
summary(tls)
```

### 3.0 - Analysis 

#### Thin point cloud

```{r thin}
tls <- tlsSample(tls, smp.voxelize(0.02))
```

#### Extract `treeMap` (pancakes)

```{r}
mh <- map.hough( 
  min_h = 0.5,   # min height
  max_h = 2,     # max height
  h_step = 0.5) # interval 
map <- treeMap(tls, method = mh)
```

#### Classify `treePoints` (tree regions)

```{r}
tls <- treePoints(tls, map, trp.crop(l = 0.5, circle = TRUE))
```

#### Classify `stemPoints`

```{r, warning=FALSE, message=FALSE}
sth <- stm.hough(
  h_step = 0.5,
  max_d = 0.75,
  h_base = c(0.5, 2))
tls <- stemPoints(tls, sth)
```

#### Generate `tlsInventory` (height and dbh)

```{r}
dbh <- shapeFit(shape = 'circle', algorithm='ransac', n=20) 
inv_cir <- tlsInventory(tls, dw=1, hp=.999, d_method=dbh)
```

### 4.0 - Clean data 

```{r}
# Calculate diameter from radius
inv_cir[, 'DBH'] = inv_cir$Radius*2
# Drop na values
df.na <- na.omit(inv_cir)
# Create a subset
inv_df <- subset(df.na, select = c(X, Y, H, DBH))
```

### 5.0 - Export to CSV 

```{r}
export_name <- paste(dir_out, 'xyh_dbh.csv', sep = '')
write.csv(inv_df, export_name)
```

These results can now be used as is or fed into the next analysis workbook. 

## Contact

**Justinn J. Jones '24, GISP**

Graduate Research Assistant / FAA sUAS RP

LASERS Lab / WFES 360

Department of Ecology & Conservation Biology

College of Agriculture & Life Sciences

[justinn.j.jones\@tamu.edu](mailto:justinn.j.jones@tamu.edu)

**TEXAS A&M UNIVERSITY**
