---
title: "#1: Measuring Trees - R notebook"
# author: "Justinn J. Jones '24, GISP"
# date: "`r Sys.Date()`"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Howdy!

This R markdown notebook is the first step towards automating the process of measuring trees.

This notebook assumes the RMD resides directly in the "main directory" with a '\_input' folder housing the input files. We will create and use an 'output' folder for output files.

### 1.0 - Setup Notebook

#### Load packages

```{r load, echo = T, results = 'hide', warning=FALSE, message=FALSE}
library(lidR)
library(raster)
library(sf)
library(ggplot2)
library(terra)
library(mapview)
```

#### Setup the workspace

Verify the working directory.

```{r set wd}
wd <- getwd()
wd
```

Set input directory and list the files.

```{r input dir}
dir_input <- './_input'
list.files(dir_input)
```

Name the objects for later use.

```{r input vars}
las_in <- "lidar_point_cloud.las"
chm_in <- "CHM_50cm.tif"
```

Create the 'output' directory.

```{r output, results = 'hide'}
dir_out <- paste(wd, '/output', sep = '')
ifelse(!dir.exists(file.path(dir_out)), dir.create(file.path(dir_out)), FALSE)
```

### 2.0 - Load Data

Read in the lidar point cloud and generate a summary of the dataset.

```{r load las, warning=FALSE}
las <- readLAS(paste(dir_input, '/', las_in, sep = ''))
summary(las)
```

Read in Canopy Height Model (CHM) and view a summary of the data.

```{r load chm}
chm1 <- raster(paste(dir_input, '/', chm_in, sep = ''))
chm1
```

View CHM.

```{r view chm}
chm_rast <- rast(chm1)
plot(chm_rast)
```

### 3.0 - Analysis

#### Generate Tree Top points

This tool uses an auto-adjusting window size to find tree tops with a local maximum filter. The variable window size is based on tree height and crown width. These variables will be set manually.

NOTE: The variable window size function does not allow for negative or null values, so the lidar point cloud minimum values must be set to Z=0 and minimum y-intercept to 0.01 for it to run without error.

Plot the 'y=mx+b' for tree height vs crown width for verification. These values were calculated from a sub-sample of trees in the dataset.

```{r plot line}
slope = 1.15
inter = -0.58

ggplot(data=data.frame( x=c(-1,10),y=c(-1,10) ), aes(x=x,y=y)) + 
  geom_blank() +
  geom_abline(slope=slope, intercept=inter, col="red", linewidth=2) +
  geom_abline(slope=slope, intercept=0, col="blue", linewidth=1) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  labs(title = 'Tree height compared to crown width', 
       x = 'Tree Height (m)', y = 'Crown Width (m)')
```

Above, the red line represents the actual data, while the blue line has been adjusted to represent only positive values.

Set variable window size and find tree tops with a minimum height of 4 meters using a circular window.

```{r tree tops}
f1 <- function(x) {x * slope + .01}
ttops1 <- locate_trees(las, lmf(ws = f1, 
                                hmin = 4, 
                                shape = c("circular")))
```

#### Generate Crown Area polygons

This tool uses a seed plus growing region algorithm to generate tree crown areas. The tree top points will be used as the seed for each tree.

```{r crown areas}
seg1 <- segment_trees(
  las, dalponte2016(chm1, ttops1, 
                    th_tree = 3.5, # Threshold below which a pixel cannot be a tree. Default is 2.
                    th_seed = 0.45, # Growing threshold 1, should be between 0 and 1, default is 0.45.
                    th_cr = 0.55, # Growing threshold 2, should be between 0 and 1, default is 0.55.
                    max_cr = 20, # Maximum value of crown diameter (in pixels), default is 10.
                    ID = "treeID"))
crowns1 <- crown_metrics(seg1, .stdtreemetrics, geom = "convex")
```

### 4.0 - Review Tree Tops and Crown Areas

Use an interactive map to verify the points and polygons.

```{r tt points}
mapview(list(ttops1, crowns1))
```

Generate a histogram of tree heights with a bin width of 0.5 meters.

```{r tt hist}
tt_hist <- ggplot(ttops1, aes(Z)) + 
  geom_histogram(binwidth = 0.5) + 
  labs(title = 'Histogram of tree heights', x = 'Tree Height (m)')
tt_hist
```

Generate a histogram of crown areas with a bin width of 10 m\^2.

```{r cw hist}
cw_hist <- ggplot(crowns1, aes(convhull_area)) + 
  geom_histogram(binwidth = 10) + 
  labs(title = 'Histogram of crown areas', x = 'Crown Area (m^2)')
cw_hist
```

### 5.0 - Export the results as shapefiles.

We will use 'st_zm' to drop Z and M value, but Z values are still recorded in the attribute table.

```{r export, results = 'hide', warning=FALSE, message=FALSE}
setwd(dir_out)
st_write((st_zm(ttops1, drop = T, what = "ZM")), "ttops1.shp", append = F)
st_write(crowns1, "crowns1.shp", append = F)
```

These results can now be used as is or fed into the next analysis workbook.

## Contact

**Justinn J. Jones '24, GISP**

Graduate Research Assistant / FAA sUAS RP

LASERS Lab / WFES 360

Department of Ecology & Conservation Biology

College of Agriculture & Life Sciences

justinn.j.jones\@tamu.edu

**TEXAS A&M UNIVERSITY**
